---
title: "Penguins size comparisons"
author: "Lauren Skube"
date: "11/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1. T-tests 

Mass comparison: male and female chinstrap penguins 

```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")

# Histograms 
ggplot(data = chinstraps, aes(x = body_mass_g)) +
  geom_histogram(bins = 10) +
  facet_wrap(~sex)

## QQ plot
ggplot(data = chinstraps, aes(sample = body_mass_g)) + 
  geom_qq() + 
  facet_wrap(~sex)

## Get descriptive statistics 
chinstraps %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g), 
    median_mass = median(body_mass_g), 
    sd_mass = sd(body_mass_g), 
    sample_size = n()
  )

```


Get vectors of body mass for female and male Chinstrap penguins separately. 

```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```

### T-tests, a couple different ways 

1. Is there a significant difference in body mass between male and female chinstrap penguins? 

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest

cohen.d(chinstraps_f, chinstraps_m)
```
Are male chinstrap penguins significantly larger (body mass) than female chinstrap penguins? 

```{r}
t.test(chinstraps_m, chinstraps_f, alternative = "greater")
```
## One-way ANOVA 

### 1. Table of counts by penguin species and sex 

```{r}
penguin_counts <- penguins %>% 
  count(species, sex)

penguin_counts
```

Create a subset of only female penguins 

```{r}
penguin_f <- penguins %>% 
  filter(sex == "female")

ggplot(data = penguin_f, aes(x = flipper_length_mm)) +
  geom_histogram(aes(fill = species)) + 
  facet_wrap(~species, scales = "free")

ggplot(data = penguin_f, aes(sample = flipper_length_mm)) +
  geom_qq(aes(color = species)) + 
  facet_wrap(~species, scales = "free")
```

### Find summary statistics 

Starting from penguins_f, create a summary table that groups by species, and returns the following: 

- mean of flipper length 
- sd ""
- sample size ""
- variance 'var'
- standard error = sd/sqrt(n)

```{r}
penguin_f_summary <- penguin_f %>% 
  group_by(species) %>% 
  summarize(
    mean_flipper_length = mean(flipper_length_mm),
    sd_flipper_length = sd(flipper_length_mm), 
    sample_size = n(), 
    variance_flipper_length = var(flipper_length_mm), 
    standard_error = sd(flipper_length_mm)/sqrt(n())
  )

leveneTest(flipper_length_mm ~ species, data = penguin_f)
```
Levene test p-value = 0.06, so retain the null hypothesis that the variances within groups are close enough. 

## Beeswarm plot with mean +- 1sd

```{r}
ggplot() +
  geom_beeswarm(data = penguin_f, 
                aes(x = species, y = flipper_length_mm),
                size = 2, 
                color = "goldenrod4") +
  geom_point(data = penguin_f_summary, aes(x= species, y = mean_flipper_length, size = 2)) + 
  geom_errorbar(data = penguin_f_summary, aes(x = species,
                                              ymax = mean_flipper_length + sd_flipper_length,
                                              ymin = mean_flipper_length - sd_flipper_length),
                width = 0.1) +
  theme_light()
```












